\normalfont \normalsize \textsc{University of Stuttgart} \\ [25pt] \horrule{0.5pt} \\[0.4cm] \huge Hauptseminar:\\The Logjam Attack \\ \horrule{2pt} \\[0.5cm]

### 2017-06-22 Ausarbeitung/paper.tex
grammar correction done: v1.1

Section 2 provide a formal definition of the key exchange problem using asymmetric keys and describes the solution according to \citep{diffie1976new}. In order to break Diffie-Hellman, \citep{gordon1993discrete} first proposed how the Number Field Sieve algorithm can be used to compute logarithms in prime fields. This method is rather complicated and covers a large part of this work in Section 3. While most methods presented here is about cryptography, Section 4 presents the Transport Layer Security, specified in \citep{dierks2008transport}, which is a security protocol used in the transport layer from the OSI model. This model is the most commonly used one that provide a decomposition of a network system into layers, from physical to application level. Section 5 uses the methods from the previous sections to explain Logjam in short. Section 6 presents an attack proposed by \citep{van1996diffie} that exploits a specific kind of bad parameter choices made by system designers when Diffie-Hellman key exchange is used. The second result presented in \citep{Adrian:2015:IFS:2810103.2813707} is an estimate on the cost of computing logs of group elements using 1024 bits primes. I will briefly summarize their assumptions for the estimate and their results in Section 7. A conclusion is drawn about what is necessary to make computer networks more secure in the future in Section 8.
%----------------------------------------------------------------------------------------
%	Diffie-Hellman Key Exchange
%----------------------------------------------------------------------------------------
\section{Diffie-Hellman Key Exchange}
\label{sec:DHKeyExchange}
\citep{diffie1976new} was the first paper presenting a practical solution to the privacy and the authentication problem that does not rely on secure communication channels. These kinds of channels are often not given or so inefficient to use that the benefit of telecommunication is nullified. Hence, it was a historical and great advance in computer networking and cryptography. They presented an asymmetric approach with trap door functions that have an exponential cipher-cryptanalyst ratio. Because of this, a security system can choose feasible large private keys that are at the same time infeasible to infer from public keys and encrypted messages w.r.t. computation time.
\subsection{Problem Formulation}
Given:
\begin{itemize}
	\item $A, B$, the communication partners.
	\item Only insecure communication channels are available.
	\item $\{M\}$, a finite message space which contain all possible messages $M$ to be exchanged.
\end{itemize}
Determine:
\begin{itemize}
	\item $C = (\{E_K\}_{K\in\{K\}}, \{D_K\}_{K\in\{K\}})$, a \textit{public key cryptosystem} with mutual inverse functions
	\begin{itemize}
		\item $E_K:\{M\}\rightarrow\{M\}$
		\item $D_K:\{M\}\rightarrow\{M\}$
	\end{itemize}
	denoted as \textit{encryption} and \textit{decryption transformation} respectively.
	\item $\{K\}$, a suitable set of keys.
	\item Such that
	\begin{enumerate}
	\item $\forall K\in\{K\}:M = D_K(E_K(M))$
	\item $\forall K\in\{K\},M\in\{M\}:E_K(M), D_K(M)$ are easy to compute.
	\item $\forall^\infty K\in\{K\},f:\{E_K\}_{K\in\{K\}}\rightarrow \{D_K\}_{K\in\{K\}}:D_K = f(E_K)$ is infeasible to compute.
	\item $\forall K\in\{K\}\exists g:\{K\}\rightarrow(\{E_K\}_{K\in\{K\}}, \{D_K\}_{K\in\{K\}}):g(K) = (E_K,D_K)$ is feasible to compute.
	\end{enumerate}
\end{itemize}
The goal is therefore to find a key space $\{K\}$ and one transformation generator $g$ to generate public and private keys $(E_K,D_K)$ that fulfil these four conditions above.

@@

Now if $A$ and $B$ want to exchange a session key, they first choose public and private keys $E_A$, $E_B$, $D_A$, $D_B$ from a number field $\mathrm{GF}(q)$ with a sufficient size $q$ and send each other their public keys over the insecure communication channel. Then, they can calculate the common session key
\begin{eqnarray}
K_{A,B} & = & \alpha^{D_AD_B} \mod q\\
& = & (\alpha^{D_A})^{D_B} \mod q\\
& = & {E_A}^{D_B} \mod q\label{eqn:keyA}\\
& = & (\alpha^{D_B})^{D_A} \mod q\\
& = & {E_B}^{D_A} \mod q.\label{eqn:keyB}
\end{eqnarray}
While $A$ calculates the right term in \ref{eqn:keyA}, $B$ calculates the right term in \ref{eqn:keyB} and both of them have now the same session key $K_{A,B}$.

@@

\paragraph{Cryptanalysis}
The attacker can compute $K_{A,B}$ either by computing $D_A$ from $E_B$ or $D_B$ from $E_A$, both are log-problems in finite fields. Thus, the attacker is successful iff he or she can effort exponentially more computation power that $A$ and $B$.

@@

\subsubsection{Smoothness Assumption}
Define that an integer number $x$ is $B$-smooth, if all prime factors of $x$ are smaller than $B$. NFS assumes the base $\alpha$ to be $B$-smooth where $B$ is an upper bound for the prime values in the factor base (which is the output of the precomputation). $B$ is restricted by our computation capabilities. If $\alpha$ is not $B$-smooth, we can use a $B$-smooth $\alpha'$ instead. Then, we can compute $\log_{\alpha'}\alpha$ and transform back to the original problem with
\begin{equation}
\log_\alpha E\equiv\frac{\log_{\alpha'}E}{\log_{\alpha'}\alpha}\mod (q-1)
\text{ .}
\end{equation}
If $\alpha'$ is not a generator for $\mathrm{GF}(q)^*$, the logs on the right hand side might not exist. $\alpha'$ can be checked for feasibility by factoring $(q-1)$ with the NFS factoring algorithm and check
\begin{equation}
(a')^{\frac{q-1}{p}} \not\equiv 1 \mod q,\quad\forall p\parallel q
\text{ ,}
\end{equation}
where $p\parallel q$ denotes that $p$ is a prime divisor of $q$. The reason why $\alpha$ or $\alpha'$ needs to be $B$-Smooth is that the factor base needs to have at least one inhomogeneous relation for the logs of the factor base, using the equation
\begin{eqnarray}
\label{eqn:assumption}
\log_\alpha\alpha = 1 \equiv \sum_{p^t\parallel a}t\log_\alpha p \mod (q-1)
\text{ ,}
\end{eqnarray}
which is not guaranteed to be the case if not all log factors are considered.

@@

Next, let $\gamma\in\mathbb{C}$ be the root of $f$, define $K = \mathbb{Q}(\gamma) = \mathbb{Q}\cup\{\gamma\}$ and $O_K$ the ring of integers in $K$. Since $\mathrm{gcd}(q,\Delta_f) = 1$, $\hat{q}:=(q,\gamma-m)$ is a first-degree prime factor of $q\in O_K$. It follows that $O_K/\hat{q}\cong\mathrm{GF}(q)$ and we can define a homomorphism $\phi:\mathbb{Z}[\alpha]\rightarrow\mathbb{Z}/q\mathbb{Z}$, providing us the representation we aimed for. We denote a prime ideal $s\in O_K$ as \textit{good} if $s$ does not divide the index $[O_K:\mathbb{Z}[\gamma]]$, otherwise \textit{bad}.

@@

\subsubsection{Linear Algebra}
For each $A \in \mathcal{A},\ A = \mathbb{Z}^{S\times T},\ S > T$, where each row has at most $X$ non-zero entries, find a linear relation over $\mathbb{Q}$ for the rows of $A$. This is accomplished in four steps, shown for the case $S=T+1$. This procedure was developed by Carl Pomerance (*1944) which was not separately published, instead only for helping out in \citep{gordon1993discrete}.
\paragraph{Step 1}Attempt to compute the rank $r = \mathrm{rk}(A)$. For that, choose a random prime $q_0\leq XT\log T$. Later, we will check if that choice was feasible. If not, come back here and sample another one. Then, use Gaussian elimination mod $q_0$ to find the rank $r_0$ of $A\mod q_0$. Rearrange the rows so that the first $r_0$ rows are linearly independent mod $q_0$. Now the rearranged matrix is
\begin{equation}
A' =
\begin{pmatrix}
v_1\\v_2\\\vdots\\v_{T\!+\!1}
\end{pmatrix}
=
\begin{pmatrix}
a'_{11} & a'_{12} & \cdots & a'_{1T} \\
a'_{21} & a'_{22} & \cdots & a'_{2T} \\
\vdots & \vdots & \ddots & \vdots \\
a'_{T\!+\!11} & a'_{T\!+\!12} & \cdots & a'_{T\!+\!1T} \\
\end{pmatrix}
\text{ .}
\end{equation}
The result of the Gaussian elimination provide a submatrix $\hat{A} = \mathbb{Z}^{r_0\times r_0}$ of the first $r_0$ rows and columns that is nonsingular mod $q_0$,
\begin{equation}
\hat{A} =
\begin{pmatrix}
v_1\\v_2\\\vdots\\v_{r_0}
\end{pmatrix}
=
\begin{pmatrix}
a'_{11} & a'_{12} & \cdots & a'_{1r_0} \\
a'_{21} & a'_{22} & \cdots & a'_{2r_0} \\
\vdots & \vdots & \ddots & \vdots \\
a'_{r_01} & a'_{r_02} & \cdots & a'_{r_0r_0} \\
\end{pmatrix}
\text{ .}
\end{equation}
\paragraph{Step 2} Attempt to express $v_{r_0+1}$ as a linear combination of $v_1,...,v_{r_0}\mod q$ for each $q\leq XT\log T$. For some $q$ it will be successful, for others not. Denote the product of all successful primes as $\boldsymbol{P}$ and the product of all unsuccessful primes as $\boldsymbol{P'}$. If $\boldsymbol{P'} > XT\log T$, then repeat Step 1 (sample a new random prime number). Otherwise, continue with the next step.
\paragraph{Step 3} Attempt to find the determinant $\mathrm{det}(\hat{A})$ of $\hat{A}$. For each prime $q|\boldsymbol{P}$, use the Wieldemann's probabilistic determinant algorithm \citep{wiedemann1986solving} to compute an integer $D_q\in\{0,...,q-1\}$, which is the determinant of $A\mod q$ with probability of at least $1-(XT)^{-2}$, i.e.
\begin{equation}
\mathbb{P}(D_q = \det(A\mod q)\mid A,q) \geq 1-(XT)^{-2}
\text{ .}
\end{equation}
For each $D_q$ we can write an congruence
\begin{equation}
D_0 \equiv D_q \mod q,\quad\forall q|\boldsymbol{P}
\end{equation}
and use the Chinese remainder theorem to compute $D_0$ fulfilling these congruences and being closest to zero. Repeat this step until we have found a value that lies within the bound
\begin{equation}
0 < |D_0| \leq (X^{\frac{1}{2}}T)^T
\text{ .}
\end{equation}

@@

\subsection{Individual Log Calculation}
\label{subsec:indiviualLogCalculation}
Now that we have our database ready, we can calculate a private key $D$ given $E$. The first step is to find a sort of a decomposition of $E$ by sampling random integers $l\in{1,..,q-1}$ until we have found an $l$ that satisfies
\begin{equation}
\alpha^lE\equiv p_1p_2\cdots p_t\mod q
\text{ ,}
\end{equation}
where each $p_i$ should be "small", for example $\leq q^\frac{1}{k}$. If the discrete logs $x_i$ of each $q_i$ were known, we could compute $D$ with
\begin{equation}
D = \left(\sum_{i=1}^{t}x_i\right)-l \mod q
\text{ .}
\end{equation}
Finding the log for a certain $p_i$ is a similar procedure as the precomputation, where we computed the logs of factors of $\alpha$. We need to find a suitable polynomial that can be used to express relations between $p_i$ and the factor base $\mathcal{B}'$ in form of linear equations. These equations can also be found by sieving through co-prime pairs $(c,d)$ and cancelling out factors not in $\mathcal{B}'$. This gives us the desired result where $p_i$ can be represented by factors in $\mathcal{B}'$, i.e.
\begin{equation}
p_i \equiv \prod_{s\in\mathcal{B}'}s^{z'_s} \mod q
\end{equation}
and thus
\begin{equation}
\log_{\alpha}p_i = \sum_{s\in\mathcal{B}'}z'_s\log_{\alpha}s \mod (q-1)
\text{ ,}
\end{equation}
where $z'_s$ is some term that come from intermediate calculation steps that I neglected for simplicity. As described above, the private key $D$ can now be calculate from the $\log_{\alpha}p_i$'s.

Note that improved versions of the NFS algorithm for logarithm calculation exist, such as the one proposed in \citep{joux2003improvements}.
%----------------------------------------------------------------------------------------
%	Transport Layer Security
%----------------------------------------------------------------------------------------
\section{Transport Layer Security}
\label{sec:tls}
The document \citep{dierks2008transport} presents the full Transport Layer Security (TLS) protocol specification. I summarize the handshake section from there because the Logjam attack exploits the protocol flaw exactly here. An abbreviated message flow diagram is shown in Figure \ref{fig:tls}.
\begin{figure*}[h]
	\centering
	\includegraphics[width=10cm]{../Grafiken/TLS/tls}
	\caption{The message flow diagram of the TLS handshake, abbreviated.}
	\label{fig:tls}
\end{figure*}
\subsection{Hello Messages}
These messages initiate the handshake and exchange information about which cipher methods are supported on each side.
\paragraph{Hello Request} This message is optional and can be sent by the server at any time. The meaning is that the server wishes to renegotiate the connection. It expects to receive the \verb|ClientHello| message next.
\paragraph{Client Hello} With this message, the client wants to establish or re-establish a connection to the server. Several information are sent that are used for negotiation. The \verb|client_version| tells which TLS is used by the client. A random structure that consists of a time stamp and some random bytes are typically used for security checks, for example avoiding replay attacks. The client also offers a list of cipher and compression methods the server can choose from. It can use a session id to reuse security agreements from earlier or other session to abbreviate the handshake.
\paragraph{Server Hello} The server replies symmetrically by sending an \verb|server_version| which is lower or equals the \verb|client_version| and also a random structure that must not depend on the client's random structure. It selects the algorithms for cipher and compression. If none of these algorithms match, the handshake fails. If the client wished to reuse a session, the server sends the same session id back, otherwise generates a new one.
\paragraph{Server Hello Done}
This message indicates the end of a \verb|ServerHello| message, because its parameters does not have fixed length. The client should check all these parameter and also the server's certificate.

@@

\subsection{Pollard lambda} This method is a randomized one that solves a special case of the log problem $E = \alpha^D\mod q$, where the solution $D$ is known to be in range $b<D<b+w$ for some $b,w\in \mathbb{N}$. Using some random function $f:\mathbb{N}\rightarrow R$, with $R\subsetneq\mathbb{N}$ being the range we want to randomize over, ideally tuned to maximize the probability of a success, two sequences of numbers are generated:
\begin{equation}
\begin{array}{lcl}
T & = & (y_0',...,y_N'),\\
W & = & (y_0,...,y_M)
\text{ .}
\end{array}
\end{equation}
Sequence $T$ starts with the value $y_0' = g^{b+w}$ and generate the remaining elements as $y_{i+1}' = y_i'g^{f(y_i')}$. For $W$, $y_0 = E$ and $y_{i+1} = y_ig^{f(y_i)}$. The log distance between $y_0'$ and $y_N'$ that set both in the following relation
\begin{equation}
\log_\alpha(y_N') = \log_\alpha(y_0') + d_N'
\text{ ,}
\end{equation}
is
\begin{equation}
d_N' = \sum_{i=0}^{N'-1}f(y_i')\mod (q-1)
\text{ .}
\end{equation}
and for $W$ analogously. The $T$ sequence stops at some $N$, while the $W$ sequence stops if there is a hit at some point $M$ such that $y_M = y_N'$ or if $d_M$ exceeds $w+d_N'$. In the first case, the solution
\begin{equation}
D = b + w + d_N' - d_M \mod (q-1)
\end{equation}
is found. Otherwise this iteration failed and we need to restart this procedure.

\subsection{Pohlig-Hellman Decomposition} This method reduces the complexity of the log problem by reducing the size of the problem instance. The idea is to find a good way to infer the original logarithm from smaller logarithms efficiently. Then, the remaining problem is to use existing methods to compute these smaller logarithms. However, the requirement is that the prime power factorization
\begin{equation}
q-1 = \prod_{i=1}^v p_i^{c_i}
\end{equation}
is given. The procedure involve two steps. First, compute $E^{\frac{n}{p_i^{c_i}}},\ i=1..v$. Denote $E':= E^{\frac{n}{p_i^{c_i}}}$ and $\alpha' := \alpha^{\frac{n}{p_i^{c_i}}}$. Because
\begin{equation}
E' = \alpha'^D
\text{ ,}
\end{equation}
the reduced problem definition is now
\begin{equation}
E' = \alpha'^{x_i} \mod (p_i),\quad i=1..v
\text{ .}
\end{equation}
Compute, for example with Pollard lambda method, each $x_i$. With the Chinese remainder theorem we can compute $D$ with the congruences
\begin{equation}
D \equiv x_i \mod p_i
\text{ ,}
\end{equation}
because
\begin{equation}
\prod_{i=1}^v p_i^{c_i} = q-1
\text{ .}
\end{equation}

@@

\paragraph{Runtime Complexity} Let $u$ be the number of bits of the binary representation of $D$, i.e. $u = \log D$. Before this method was discovered, the best attack was known to run at $\mathcal{O}(2^{\frac{u}{2}})$ time. This method has overall cost of $\mathcal{O}(2^{\frac{u}{4}})$, which made key exchange systems that expose these properties even more insecure.

@@

Based on the most recent records in log computation and factorization at that time, the authors extrapolated a cost of 45 million core-years to do the precomputation for a group using 1024 bit primes and 30 core-days for the descent step. Assuming specialized hardware to do the job, they estimate a cost of \$8M to buy enough hardware to do the computation in one year just for Sieving and another \$11B to buy multiple Titan supercomputer clusters for the Linear Algebra step. This sum is indeed within the budget that the U.S. government planned for a cryptologic program, where the NSA is part of. Additionally, the leaked documents also contain an architecture of the decrypt infrastructure, which is perfectly consistent with NFS breaking Diffie-Hellman.

>>>>

Section 2 provides a formal definition of the key exchange problem using asymmetric keys and describes the solution according to \citep{diffie1976new}. In order to break Diffie-Hellman, \citep{gordon1993discrete} first proposed how the Number Field Sieve algorithm can be used to compute logarithms in prime fields. This method is rather complicated and covers a large part of this work in Section 3. While most methods presented here is about cryptography, Section 4 presents the Transport Layer Security, specified in \citep{dierks2008transport}, which is a security protocol used in the transport layer from the OSI model. This model is the most commonly used one that provides a decomposition of a network system into layers, from physical to application level. Section 5 uses the methods from the previous sections to explain Logjam in short. Section 6 presents an attack proposed by \citep{van1996diffie} that exploits a specific kind of bad parameter choices made by system designers when Diffie-Hellman key exchange is used. The second result presented in \citep{Adrian:2015:IFS:2810103.2813707} is an estimate on the cost of computing logs of group elements using 1024 bits primes. I will briefly summarize their assumptions for the estimate and their results in Section 7. A conclusion is drawn about what is necessary to make computer networks more secure in the future in Section 8.
%----------------------------------------------------------------------------------------
%	Diffie-Hellman Key Exchange
%----------------------------------------------------------------------------------------
\section{Diffie-Hellman Key Exchange}
\label{sec:DHKeyExchange}
\citep{diffie1976new} was the first paper presenting a practical solution to the privacy and the authentication problem that does not rely on secure communication channels. These kinds of channels are often not given or so inefficient to use that the benefit of telecommunication is nullified. Hence, it was a historical and great advance in computer networking and cryptography. They presented an asymmetric approach with trap door functions that have an exponential cipher-cryptanalyst ratio. Because of this, a security system can choose feasible large private keys that are at the same time infeasible to infer from public keys and encrypted messages w.r.t. computation time.
\subsection{Problem Formulation}
Given:
\begin{itemize}
	\item $A, B$, the communication partners.
	\item Only insecure communication channels are available.
	\item $\{M\}$, a finite message space which contains all possible messages $M$ to be exchanged.
\end{itemize}
Determine:
\begin{itemize}
	\item $C = (\{E_K\}_{K\in\{K\}}, \{D_K\}_{K\in\{K\}})$, a \textit{public key cryptosystem} with mutual inverse functions
	\begin{itemize}
		\item $E_K:\{M\}\rightarrow\{M\}$
		\item $D_K:\{M\}\rightarrow\{M\}$
	\end{itemize}
	denoted as \textit{encryption} and \textit{decryption transformation} respectively.
	\item $\{K\}$, a suitable set of keys.
	\item Such that
	\begin{enumerate}
	\item $\forall K\in\{K\}:M = D_K(E_K(M))$
	\item $\forall K\in\{K\},M\in\{M\}:E_K(M), D_K(M)$ are easy to compute.
	\item $\forall^\infty K\in\{K\},f:\{E_K\}_{K\in\{K\}}\rightarrow \{D_K\}_{K\in\{K\}}:D_K = f(E_K)$ is infeasible to compute.
	\item $\forall K\in\{K\}\exists g:\{K\}\rightarrow(\{E_K\}_{K\in\{K\}}, \{D_K\}_{K\in\{K\}}):g(K) = (E_K,D_K)$ is feasible to compute.
	\end{enumerate}
\end{itemize}
The goal is therefore to find a key space $\{K\}$ and one transformation generator $g$ to generate public and private keys $(E_K,D_K)$ that fulfil these four conditions above.

@@

Now if $A$ and $B$ want to exchange a session key, they first choose public and private keys $E_A$, $E_B$, $D_A$, $D_B$ from a number field $\mathrm{GF}(q)$ with a sufficient size $q$ and send each other their public keys over the insecure communication channel. Then, they can calculate the common session key
\begin{eqnarray}
K_{A,B} & = & \alpha^{D_AD_B} \mod q\\
& = & (\alpha^{D_A})^{D_B} \mod q\\
& = & {E_A}^{D_B} \mod q\label{eqn:keyA}\\
& = & (\alpha^{D_B})^{D_A} \mod q\\
& = & {E_B}^{D_A} \mod q.\label{eqn:keyB}
\end{eqnarray}
While $A$ calculates the right term in \ref{eqn:keyA}, $B$ calculates the right term in \ref{eqn:keyB} and both of them now have the same session key $K_{A,B}$.

@@

\paragraph{Cryptanalysis}
The attacker can compute $K_{A,B}$ either by computing $D_A$ from $E_B$ or $D_B$ from $E_A$, both are log-problems in finite fields. Thus, the attacker is successful iff he or she can afford exponentially more computation power than $A$ and $B$.

@@

\subsubsection{Smoothness Assumption}
Define that an integer number $x$ is $B$-smooth, if all prime factors of $x$ are smaller than $B$. NFS assumes the base $\alpha$ to be $B$-smooth where $B$ is an upper bound for the prime values in the factor base (which is the output of the precomputation). $B$ is restricted by our computation capabilities. If $\alpha$ is not $B$-smooth, we can use a $B$-smooth $\alpha'$ instead. Then, we can compute $\log_{\alpha'}\alpha$ and formulate our original problem with
\begin{equation}
\log_\alpha E\equiv\frac{\log_{\alpha'}E}{\log_{\alpha'}\alpha}\mod (q-1)
\text{ .}
\end{equation}
If $\alpha'$ is not a generator for $\mathrm{GF}(q)^*$, the logs on the right hand side might not exist. $\alpha'$ can be checked for feasibility by factoring $(q-1)$ with the NFS factoring algorithm and check
\begin{equation}
(a')^{\frac{q-1}{p}} \not\equiv 1 \mod q,\quad\forall p\parallel q
\text{ ,}
\end{equation}
where $p\parallel q$ denotes that $p$ is a prime divisor of $q$. The reason why $\alpha$ or $\alpha'$ needs to be $B$-Smooth is that the factor base needs to have at least one inhomogeneous relation for the logs of the factor base, using the equation
\begin{eqnarray}
\label{eqn:assumption}
\log_\alpha\alpha = 1 \equiv \sum_{p^t\parallel a}t\log_\alpha p \mod (q-1)
\text{ ,}
\end{eqnarray}
which is not guaranteed to be the case if not all log factors are considered.

@@

Next, let $\gamma\in\mathbb{C}$ be the root of $f$, define $K = \mathbb{Q}(\gamma) = \mathbb{Q}\cup\{\gamma\}$ and $O_K$ the ring of integers in $K$. Since $\mathrm{gcd}(q,\Delta_f) = 1$, $\hat{q}:=(q,\gamma-m)$ is a first-degree prime factor of $q\in O_K$. It follows that $O_K/\hat{q}\cong\mathrm{GF}(q)$ and we can define a homomorphism $\phi:\mathbb{Z}[\alpha]\rightarrow\mathbb{Z}/q\mathbb{Z}$, providing us the representation we aimed for. We denote a prime ideal $s\in O_K$ as \textit{good} if $s$ does not divide the index $[O_K:\mathbb{Z}[\gamma]]$, otherwise we denote it as \textit{bad}.

@@

\subsubsection{Linear Algebra}
For each $A \in \mathcal{A},\ A = \mathbb{Z}^{S\times T},\ S > T$, where each row has at most $X$ non-zero entries, find a linear relation over $\mathbb{Q}$ for the rows of $A$. This is accomplished in four steps, shown for the case $S=T+1$. This procedure was developed by Carl Pomerance (*1944) which was not separately published, instead only for helping out in \citep{gordon1993discrete}.
\paragraph{Step 1}Attempt to compute the rank $r = \mathrm{rk}(A)$. For that, choose a random prime $q_0\leq XT\log T$. Later, we will check if that choice was feasible. If not, come back here and sample another one. Then, use Gaussian elimination mod $q_0$ to find the rank $r_0$ of $A\mod q_0$. Rearrange the rows so that the first $r_0$ rows are linearly independent mod $q_0$. Now the rearranged matrix is
\begin{equation}
A' =
\begin{pmatrix}
v_1\\v_2\\\vdots\\v_{T\!+\!1}
\end{pmatrix}
=
\begin{pmatrix}
a'_{11} & a'_{12} & \cdots & a'_{1T} \\
a'_{21} & a'_{22} & \cdots & a'_{2T} \\
\vdots & \vdots & \ddots & \vdots \\
a'_{T\!+\!11} & a'_{T\!+\!12} & \cdots & a'_{T\!+\!1T} \\
\end{pmatrix}
\text{ .}
\end{equation}
The result of the Gaussian elimination provides a submatrix $\hat{A} = \mathbb{Z}^{r_0\times r_0}$ of the first $r_0$ rows and columns that is nonsingular mod $q_0$,
\begin{equation}
\hat{A} =
\begin{pmatrix}
v_1\\v_2\\\vdots\\v_{r_0}
\end{pmatrix}
=
\begin{pmatrix}
a'_{11} & a'_{12} & \cdots & a'_{1r_0} \\
a'_{21} & a'_{22} & \cdots & a'_{2r_0} \\
\vdots & \vdots & \ddots & \vdots \\
a'_{r_01} & a'_{r_02} & \cdots & a'_{r_0r_0} \\
\end{pmatrix}
\text{ .}
\end{equation}
\paragraph{Step 2} Attempt to express $v_{r_0+1}$ as a linear combination of $v_1,...,v_{r_0}\mod q$ for each $q\leq XT\log T$. For some $q$ it will be successful, for others not. Denote the product of all successful primes as $\boldsymbol{P}$ and the product of all unsuccessful primes as $\boldsymbol{P'}$. If $\boldsymbol{P'} > XT\log T$, then repeat Step 1 (sample a new random prime number). Otherwise, continue with the next step.
\paragraph{Step 3} Attempt to find the determinant $\mathrm{det}(\hat{A})$ of $\hat{A}$. For each prime $q|\boldsymbol{P}$, use the Wieldemann's probabilistic determinant algorithm \citep{wiedemann1986solving} to compute an integer $D_q\in\{0,...,q-1\}$, which is the determinant of $A\mod q$ with probability of at least $1-(XT)^{-2}$, i.e.
\begin{equation}
\mathbb{P}(D_q = \det(A\mod q)\mid A,q) \geq 1-(XT)^{-2}
\text{ .}
\end{equation}
For each $D_q$ we can write an congruence
\begin{equation}
D_0 \equiv D_q \mod q,\quad\forall q|\boldsymbol{P}
\end{equation}
and use the Chinese remainder theorem to compute $D_0$ fulfilling these congruences and being closest to zero. Repeat this step until we have found a value that lies within the bound
\begin{equation}
0 < |D_0| \leq (X^{\frac{1}{2}}T)^T
\text{ .}
\end{equation}

@@

\subsection{Individual Log Calculation}
\label{subsec:indiviualLogCalculation}
Now that we have our database ready, we can calculate a private key $D$ given $E$. The first step is to find a sort of a decomposition of $E$ by sampling random integers $l\in{1,..,q-1}$ until we have found an $l$ that satisfies
\begin{equation}
\alpha^lE\equiv p_1p_2\cdots p_t\mod q
\text{ ,}
\end{equation}
where each $p_i$ should be "small", for example $\leq q^\frac{1}{k}$. If the discrete logs $x_i$ of each $q_i$ were known, we could compute $D$ with
\begin{equation}
D = \left(\sum_{i=1}^{t}x_i\right)-l \mod q
\text{ .}
\end{equation}
Finding the log for a certain $p_i$ is a similar procedure as the precomputation, where we computed the logs of factors of $\alpha$. We need to find a suitable polynomial that can be used to express relations between $p_i$ and the factor base $\mathcal{B}'$ in form of linear equations. These equations can also be found by sieving through co-prime pairs $(c,d)$ and cancelling out factors not in $\mathcal{B}'$. This gives us the desired result where $p_i$ can be represented by factors in $\mathcal{B}'$, i.e.
\begin{equation}
p_i \equiv \prod_{s\in\mathcal{B}'}s^{z'_s} \mod q
\end{equation}
and thus
\begin{equation}
\log_{\alpha}p_i = \sum_{s\in\mathcal{B}'}z'_s\log_{\alpha}s \mod (q-1)
\text{ ,}
\end{equation}
where $z'_s$ is some term that comes from intermediate calculation steps that I neglected for simplicity. As described above, the private key $D$ can now be calculated from the $\log_{\alpha}p_i$'s.

Note that improved versions of the NFS algorithm for logarithm calculation exist, such as the one proposed in \citep{joux2003improvements}.
%----------------------------------------------------------------------------------------
%	Transport Layer Security
%----------------------------------------------------------------------------------------
\section{Transport Layer Security}
\label{sec:tls}
The document \citep{dierks2008transport} presents the full Transport Layer Security (TLS) protocol specification. I summarize the handshake section from there because the Logjam attack exploits the protocol flaw exactly here. An abbreviated message flow diagram is shown in Figure \ref{fig:tls}.
\begin{figure*}[h]
	\centering
	\includegraphics[width=10cm]{../Grafiken/TLS/tls}
	\caption{The message flow diagram of the TLS handshake, abbreviated.}
	\label{fig:tls}
\end{figure*}
\subsection{Hello Messages}
These messages initiate the handshake and exchange information about which cipher methods are supported on each side.
\paragraph{Hello Request} This message is optional and can be sent by the server at any time. The meaning is that the server wishes to renegotiate the connection. It expects to receive the \verb|ClientHello| message next.
\paragraph{Client Hello} With this message, the client wants to establish or re-establish a connection to the server. Several information are sent that are used for negotiation. The \verb|client_version| tells which TLS is used by the client. A random structure that consists of a time stamp and some random bytes are typically used for security checks, for example avoiding replay attacks. The client also offers a list of cipher and compression methods the server can choose from. It can use a session id to reuse security agreements from earlier or other session to abbreviate the handshake.
\paragraph{Server Hello} The server replies symmetrically by sending an \verb|server_version| which is lower or equals the \verb|client_version| and also a random structure that must not depend on the client's random structure. It selects the algorithms for cipher and compression. If none of these algorithms match, the handshake fails. If the client wished to reuse a session, the server sends the same session id back, otherwise it generates a new one.
\paragraph{Server Hello Done}
This message indicates the end of a \verb|ServerHello| message, because its parameters do not have fixed lengths. The client should check all these parameters and also the server's certificate.

@@

\subsection{Pollard lambda} This method is a randomized one that solves a special case of the log problem $E = \alpha^D\mod q$, where the solution $D$ is known to be in range $b<D<b+w$ for some $b,w\in \mathbb{N}$. Using some random function $f:\mathbb{N}\rightarrow R$, with $R\subsetneq\mathbb{N}$ being the range we want to randomize over, ideally tuned to maximize the probability of a success, two sequences of numbers are generated:
\begin{equation}
\begin{array}{lcl}
T & = & (y_0',...,y_N'),\\
W & = & (y_0,...,y_M)
\text{ .}
\end{array}
\end{equation}
Sequence $T$ starts with the value $y_0' = g^{b+w}$ and generates the remaining elements as $y_{i+1}' = y_i'g^{f(y_i')}$. For $W$, $y_0 = E$ and $y_{i+1} = y_ig^{f(y_i)}$. The log distance between $y_0'$ and $y_N'$ that sets both in the following relation
\begin{equation}
\log_\alpha(y_N') = \log_\alpha(y_0') + d_N'
\text{ ,}
\end{equation}
is
\begin{equation}
d_N' = \sum_{i=0}^{N'-1}f(y_i')\mod (q-1)
\text{ .}
\end{equation}
and for $W$ analogously. The $T$ sequence stops at some $N$, while the $W$ sequence stops if there is a hit at some point $M$ such that $y_M = y_N'$ or if $d_M$ exceeds $w+d_N'$. In the first case, the solution
\begin{equation}
D = b + w + d_N' - d_M \mod (q-1)
\end{equation}
is found. Otherwise this iteration failed and we need to restart this procedure.

\subsection{Pohlig-Hellman Decomposition} This method reduces the complexity of the log problem by reducing the size of the problem instance. The idea is to find a good way to infer the original logarithm from smaller logarithms efficiently. Then, the remaining problem is to use existing methods to compute these smaller logarithms. However, the requirement is that the prime power factorization
\begin{equation}
q-1 = \prod_{i=1}^v p_i^{c_i}
\end{equation}
is given. The procedure involves two steps. First, compute $E^{\frac{n}{p_i^{c_i}}},\ i=1..v$. Denote $E':= E^{\frac{n}{p_i^{c_i}}}$ and $\alpha' := \alpha^{\frac{n}{p_i^{c_i}}}$. Because
\begin{equation}
E' = \alpha'^D
\text{ ,}
\end{equation}
the reduced problem definition is now
\begin{equation}
E' = \alpha'^{x_i} \mod (p_i),\quad i=1..v
\text{ .}
\end{equation}
Compute, for example with Pollard lambda method, each $x_i$. With the Chinese remainder theorem we can compute $D$ with the congruences
\begin{equation}
D \equiv x_i \mod p_i
\text{ ,}
\end{equation}
because
\begin{equation}
\prod_{i=1}^v p_i^{c_i} = q-1
\text{ .}
\end{equation}

@@

\paragraph{Runtime Complexity} Let $u$ be the number of bits of the binary representation of $D$, i.e. $u = \log D$. Before this method was discovered, the best attack was known to run at $\mathcal{O}(2^{\frac{u}{2}})$ time. This method has an overall cost of $\mathcal{O}(2^{\frac{u}{4}})$, which made key exchange systems that expose these properties even more insecure.

@@

Based on the most recent records in log computation and factorization at that time, the authors extrapolated a cost of 45 million core-years to do the precomputation for a group using 1024 bit primes and 30 core-days for individual log computation. Assuming specialized hardware to do the job, they estimate a cost of \$8M to buy enough hardware to do the computation in one year just for Sieving and another \$11B to buy multiple Titan supercomputer clusters for the Linear Algebra step. This sum is indeed within the budget that the U.S. government planned for a cryptologic program, where the NSA is part of. Additionally, the leaked documents also contain an architecture of the decrypt infrastructure, which is perfectly consistent with NFS breaking Diffie-Hellman.
